---
title: "Lab 5"
author: "Conner Smith"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
```

# Confidence Intervals

Calculate confidence interval for single sample uting the `t.test` function 

## Example 1

Create a mock sample of observation values and find the confidence interval. Then we'll learn how to report that.

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)
```

Degrees of freedom is sample size - 1 

### Risky and Tedious Way:

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g.

### A high initial investment, but safer and better way 

In-line code referencing to stored objects 

Use single backtick on either side of lower case "r" to create in-line code.

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` grams $\pm$ `r round(pika_sd, 1)` grams (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round(pika_ci$conf.int[1], 2)`, `r round(pika_ci$conf.int[2], 2)`] grams.

# Two-sample t-test 

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means =  0). 

- H0: Means difference = 0
- HA: Means difference is NOT = 0

## Example 1

```{r}
oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is  there a significant difference in mean heights for oak and pine trees?

Always need to ask questions on how representative a sample actually is. 

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r trees_t$p.value` means that there is a `r (trees_t$p.value)*100` percent chance of finding sample means that are at least as different from two other random samples if they were drawn from populations with the same mean height. 

This means we would most likely reject the null hypothesis that these samples were drawn from population with the same mean when operating with a 95% confidence interval. 

## Example 2

Uging the `mpg` dataset to compare the city gas mileage for SUVs and compact cars.

```{r}
compact_suv <- mpg %>% 
  filter(class %in% c("compact", "suv"))

```

```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram(bins = 12, aes(fill = class)) +
  facet_wrap(~ class)

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq(aes(color = class)) +
  facet_wrap(~ class)
```


```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = mean(cty),
            sd_city = sd(cty),
            sample_size = n())

car_stats
```

```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)
```

This means we would most likely reject the null hypothesis that these samples were drawn from population with the same mean when operating with a 95% confidence interval. 

Mean gas mileage for compact cars and SUVs differs significantly (t(`r cars_t$parameter`) = `r cars_t$statistic`, p < 0.001, $\alpha$ = 0.05).
